#upload2001datasets from nomis site
#source:https://www.nomisweb.co.uk/sources/census_2001_ks
#source:https://www.nomisweb.co.uk/sources/census_2001_uv

# Merge datasets by common fields to output area

```{r, comment=NA}
#cleans headingsmergeddatasets
x2001_dataset<-clean_names(x2001_dataset)
```


#2001 dataset variable selection
```{r}

x2001_oa_var <- x2001_dataset %>%

#employment
mutate(employed = economically_active- economically_active_full_time_student) %>%
rename(unemployed = economically_active_unemployed ) %>%
mutate(students= economically_active_full_time_student+
economically_inactive_student_including_full_time_students) %>%
rename(retired=economically_inactive_retired) %>%

#ethnicity
rename(white=white) %>%
rename(mixed=mixed) %>%
rename(indian=asian_asian_british_indian) %>%
rename(pakisthani=asian_asian_british_pakistani) %>%
rename(bangladeshi=asian_asian_british_bangladeshi) %>%
rename(chinese=chinese_other ) %>%
rename(black_caribbean=black_black_british_black_caribbean) %>%
rename(black_african=black_black_british_black_african) %>%

#part time
#mutate(other_ethinicity=ethnic_group_asian_asian_british_or_asian_welsh_other_asian+ethnic_group_other_ethnic_group) %>%
#select(white, mixed,indian,pakisthani,bangladeshi,chinese,black,other_ethinicity) %>%

#age
mutate('0_to_19' = age_0_to_4 + age_5_to_7 + age_8_to_9+age_10_to_14+ age_15 + age_16_to_17 + age_18_to_19 )
mutate('20_to_29'= age_20_to_24 + age_25_to_29 ) %>%
rename('30_to_44' =age_30_to_44 ) %>%
mutate('45_to_64'= age_45_to_59 + age_60_to_64 ) %>%
mutate(over65= age_65_to_74 +age_75_to_84 +age_85_to_89 +age_90_and_over) %>%

#country of birth
mutate(uk=united_kingdom_x) %>%
mutate('other_erouope'=other_western_europe_x+ eastern_europe_x ) %>%
rename(africa=africa_x) %>%
rename(asia=asia_x) %>%
mutate(america=north_america_x +south_america_x) %>%
rename(ocenia= oceania_x) %>%


#tenure
mutate(owned= owned_outright+owned_with_a_mortgage_or_loan+shared_ownership) %>%
rename('prived_rented'= rented_from_a_private_landlord_or_letting_agency) %>%
mutate('social_rented'=rented_from_council_local_authority+ rented_from_a_housing_association_registered_social_landlord )


#mode travel
mutate('work_home'=work_mainly_at_or_from_home) %>%
mutate('public_transport'=underground_metro_light_rail_tram +train +bus_minibus_or_coach ) %>%
mutate('private_transport'=driving_a_car_or_van +passenger_in_a_car_or_van ) %>%
mutate('close_distance'=motorcycle_scooter_or_moped +bicycle +on_foot ) %>%
rename('other_mode'=other_y ) %>%


#occupation
mutate('managerial_and_professional' = x1_managers_and_senior_officials + x2_professional_occupations)
mutate('skilled_non_manual' = x3_associate_professional_and_technical_occupations + x4_administrative_and_secretarial_occupations)
rename('skilled_manual' = x5_skilled_trades_occupations) %>%
mutate('partly_skilled' =x6_personal_service_occupations+x7_sales_and_customer_service_occupations+x8_process_plant_and_machine_operatives
+x9_elementary_occupations) %>%


#education
rename('no_qualification'= no_qualifications) %>%
rename('level_1'=highest_level_of_qualification_level_1_qualifications) %>%
rename('level_2'=highest_level_of_qualification_level_2_qualifications) %>%
mutate('level_3'=highest_level_of_qualification_level_3_qualifications) %>%
mutate('level_4'= highest_level_of_qualification_level_4_qualifications_and_above+highest_level_of_qualification_other_qualifications)


#Dwelling
rename(detached= whole_house_or_bungalow_detatched) %>%
rename('semi_detached'= whole_house_or_bungalow_semi_detatched) %>%
rename(terrace= whole_house_or_bungalow_terraced ) %>%
mutate(flats=flat_maisonette_or_apartment_in_a_purpose_built_block_of_flats_or_tenement ) %>%
mutate('other_accomodation'= caravan_or_other_mobile_or_temporary_structure) %>%


#work_hours
mutate('0-15'= males_part_time_1_to_5_hours+ males_part_time_6_to_15_hours + females_part_time_1_to_5_hours
mutate('16-30'=males_part_time_16_to_30_hours + females_part_time_16_to_30_hours ) %>%
mutate('31-48'=males_full_time_31_to_37_hours + males_full_time_38_to_48_hours +females_full_time_31_to_37_hours
mutate(above48=males_full_time_49_or_more_hours + females_full_time_49_or_more_hours ) %>%


#migration
rename('migrate_within_uk'= moved_into_the_area_from_within_the_uk_4) %>%
rename('migrate_outside_uk'=moved_into_the_area_from_outside_the_uk_5) %>%

#central heating
mutate(house_with_central_heating=with_central_heating_and_sole_use_of_bath_shower_and_toilet_x+
with_central_heating_without_sole_use_of_bath_shower_and_toilet_x) %>%
mutate(house_without_central_heating= without_central_heating_or_sole_use_of_bath_shower_and_toilet_x
without_central_heating_with_sole_use_of_bath_shower_and_toilet_x) %>%

#household composition
mutate('one_person' = one_person_household) %>%
mutate('family_without_children' =one_family_and_no_others_married_couple_households_no_children ) %>%
mutate('family_non_depended_children'= one_family_and_no_others_married_couple_households_all_children_non_dependent
one_family_and_no_others_lone_parent_households_all_children_non_dependent +
one_family_and_no_others_cohabiting_couple_households_all_children_non_dependent) %>%
mutate('family_depended_children'= one_family_and_no_others_married_couple_households_with_dependent_children
one_family_and_no_others_lone_parent_households_with_dependent_children+
one_family_and_no_others_cohabiting_couple_households_with_dependent_children+
other_households_with_dependent_children ) %>%

#gender
mutate(male= males_x) %>%
rename ( female= females_y ) %>%

#industry
mutate('agriculture_energy_and_water'= a_agriculture_hunting_forestry+ b_fishing+e_electricity_gas_and_water_supply
rename(manufacturing= d_manufacturing) %>%
mutate('construction'= f_constructiion ) %>%
mutate(distribution_hotels_and_restaurants= h_hotels_and_restaurants ) %>%
mutate(transport_and_communication= i_transport_storage_and_communications +g_wholesale_and_retail_trade_repair_of_motor_vehicles)
mutate(finance_business_admin= j_financial_intermediation+ k_real_estate_renting_and_business_activities
mutate(educatoin_health=m_education+n_health_and_social_work ) %>%
rename( other= o_p_q_other ) %>%

#rename colunms- rate calculation purpose
rename(geo_code=x2001_output_area) %>%
rename(total_pop=all_categories_country_of_birth_x) %>%
rename(total_accomadation=all_categories_tenure ) %>%
rename(total_16_17age= all_usual_residents_aged_16_to_74) %>%
rename(total_household= all_households_y) %>%
rename(total_industry= all_categories_occupation) %>%
rename(total_people_gender= all_people) %>%
rename(migrants= people_who_are_migrants) %>%
mutate(total_dwelling= all_household_spaces_with_residents+ all_household_spaces_with_no_residents_vacant )

#select finalvariables

dplyr::select(geo_code,total_pop,total_accomadation,total_dwelling, total_16_17age,total_household,total_industry,total_people_gender,migrants,
unemployed,employed,students,retired,white,mixed,indian,pakisthani,bangladeshi,chinese,black_african,black_caribbean,
'0_to_19','20_to_29','30_to_44','45_to_64',over65,owned,prived_rented,social_rented,
work_home,public_transport,private_transport,close_distance,other_mode,
managerial_and_professional,skilled_non_manual,skilled_manual,partly_skilled,no_qualification,level_1,level_2,level_3,level_4,
detached,semi_detached,terrace,flats,other_accomodation,'0-15','16-30','31-48',above48,uk,other_erouope,africa,asia,america,
ocenia,migrate_within_uk,migrate_outside_uk,house_with_central_heating,house_without_central_heating,one_person,family_without_children,
family_depended_children,family_non_depended_children,male,female,agriculture_energy_and_water, manufacturing,construction,
distribution_hotels_and_restaurants,transport_and_communication,finance_business_admin,educatoin_health)

```


#savedatasetselectedvariables
```{r}
x2001_oa_var %>%
saveRDS("/home/myk9/diss/x2001_oa_var.rds")
x2001_oa_var<
readRDS("/home/myk9/diss/x2001_oa_var.rds")
```


#2011dataset
#uploaddatasets2021
#sources-https://www.nomisweb.co.uk/census/2011/bulk

#clean and mearge data

#select varibles
```{r}
x2011_oa_var <- data2011_oa %>%

#employment
mutate(employed = KS601EW0016+KS601EW0017+KS601EW0018+ KS601EW0019 ) %>%
mutate(unemployed = KS601EW0019) %>%
mutate(students= KS601EW0020 + KS601EW0022) %>%
rename(retired=KS601EW0021) %>%


#ethnicity
mutate(white=KS201EW0020+KS201EW0021+KS201EW0022+KS201EW0023) %>%
mutate(mixed=KS201EW0024+KS201EW0025+KS201EW0026+KS201EW0027) %>%
rename(indian=KS201EW0028) %>%
rename(pakisthani=KS201EW0029) %>%
rename(bangladeshi=KS201EW0030) %>%
rename(chinese=KS201EW0031 ) %>%
rename(black_african=KS201EW0033) %>%
rename(black_caribbean= KS201EW0034) %>%
mutate('other_ethinicity'=KS201EW0032+KS201EW0036+KS201EW0037) %>%


#age
mutate('0_to_19' = KS102EW0020+KS102EW0021+KS102EW0022+KS102EW0023+KS102EW0024 + KS102EW0025+KS102EW0026)
mutate('20_to_29'=+KS102EW0027+KS102EW0028) %>%
rename('30_to_44' =KS102EW0029) %>%
mutate('45_to_64'=KS102EW0030+KS102EW0031 ) %>%
mutate(over65= KS102EW0032+KS102EW0033+KS102EW0034+KS102EW0035) %>%


#tenure
mutate(owned= KS402EW0010+KS402EW0011+KS402EW0012) %>%
mutate('social_rented'=KS402EW0013+KS402EW0014) %>%
mutate('prived_rented'= KS402EW0015+KS402EW0016) %>%


#mode travel
mutate('work_home'=QS701EW0002) %>%
mutate('public_transport'=QS701EW0003+QS701EW0004+QS701EW0005) %>%
mutate('private_transport'=QS701EW0006+QS701EW0008+QS701EW0009) %>%
mutate('close_distance'=QS701EW0007+QS701EW0010+QS701EW0011) %>%
rename('other_mode'=QS701EW0012) %>%


#occupation
mutate('managerial_and_professional' = KS608EW0011 + KS608EW0012) %>%
mutate('skilled_non_manual' = KS608EW0013 + KS608EW0014) %>%
rename('skilled_manual' = KS608EW0015) %>%
mutate('partly_skilled' =KS608EW0016+KS608EW0017+KS608EW0018+KS608EW0019) %>%


#education
rename('no_qualification'= KS501EW0014 ) %>%
rename('level_1'=KS501EW0015) %>%
rename('level_2'=KS501EW0016) %>%
mutate('level_3'=KS501EW0017+KS501EW0018) %>%
mutate('level_4'= KS501EW0019+KS501EW0020) %>%

#Dewlling
rename(detached= KS401EW0020) %>%
rename('semi_detached'= KS401EW0021) %>%
rename(terrace= KS401EW0022 ) %>%
mutate(flats=KS401EW0023) %>%
mutate('other_accomodation'=KS401EW0026) %>%


#work_hours
rename('0-15'= KS604EW0016) %>%
rename('16-30'=KS604EW0017) %>%
rename('31-48'=KS604EW0018) %>%
rename(above48=KS604EW0019 ) %>%


#migration
rename('migrate_within_uk'= migrants_moved_into_the_area_from_within_the_uk) %>%
rename('migrate_outside_uk'=migrants_moved_into_the_area_from_outside_the_uk) %>%

#central heating
rename(house_with_central_heating=does_have_central_heating) %>%
rename(house_without_central_heating= does_not_have_central_heating ) %>%

#household composition
rename('one_person' = one_person_household) %>%
rename('family_without_children' =one_family_only_married_or_same_sex_civil_partnership_couple_no_children )
mutate('family_non_depended_children'= one_family_only_married_or_same_sex_civil_partnership_couple_all_children_non_dependent
one_family_only_lone_parent_all_children_non_dependent +
one_family_only_cohabiting_couple_all_children_non_dependent) %>%
mutate('family_depended_children'= one_family_only_married_or_same_sex_civil_partnership_couple_dependent_children
one_family_only_cohabiting_couple_dependent_children+ one_family_only_lone_parent_dependent_children
other_household_types_with_dependent_children ) %>%

#gender
rename(male= males) %>%
rename ( female= females ) %>%

#industry
mutate('agriculture_energy_and_water'= a_agriculture_forestry_and_fishing +d_electricity_gas_steam_and_air_conditioning_supply
e_water_supply_sewerage_waste_management_and_remediation_activities) %>%
rename(manufacturing= c_manufacturing) %>%
rename('construction'= f_construction ) %>%
rename(distribution_hotels_and_restaurants= i_accommodation_and_food_service_activities ) %>%
mutate(transport_and_communication= g_wholesale_and_retail_trade_repair_of_motor_vehicles_and_motor_cycles
j_information_and_communication ) %>%
mutate(finance_business_admin= k_financial_and_insurance_activities+ l_real_estate_activities + o_public_administration_and_defence_compulsory_social_security)
mutate(educatoin_health=p_education+q_human_health_and_social_work_activities ) %>%
rename(total_dwelling= KS401EW0005) %>%

#born country
mutate(uk =europe_united_kingdom_total ) %>%
rename('other_erouope'=europe_other_europe_total ) %>%
rename(africa=africa_total) %>%
rename(asia=middle_east_and_asia_total) %>%
rename(america=the_americas_and_the_caribbean_total ) %>%
rename(ocenia= antarctica_and_oceania_total) %>%

#rename colunms for rate calculation
rename(geo_code=GeographyCode) %>%
rename(total_pop=all_usual_residents) %>%
rename(total_household= all_categories_household_composition) %>%
rename(total_industry= all_categories_industry) %>%
rename(total_employment=KS608EW0001 ) %>%
rename(migrants= migrants_total) %>%


#select variables
dplyr:: select(geo_code,total_pop, total_employment,total_industry,migrants,total_household, total_dwelling,
unemployed,employed,students,retired, white, mixed,indian,pakisthani,bangladeshi,chinese,black_african, black_caribbean,
'0_to_19','20_to_29','30_to_44','45_to_64',over65, owned,prived_rented,social_rented,
work_home,public_transport,private_transport,close_distance,other_mode,
managerial_and_professional,skilled_non_manual,skilled_manual,partly_skilled,no_qualification,level_1,level_2,level_3,level_4 ,
detached,semi_detached,terrace,flats,other_accomodation,'0-15','16-30','31-48',above48, uk,other_erouope, africa, asia, america,
ocenia, migrate_within_uk,migrate_outside_uk, house_with_central_heating, house_without_central_heating, one_person, family_without_children,
family_depended_children, family_non_depended_children, male, female, agriculture_energy_and_water, manufacturing,construction,
distribution_hotels_and_restaurants, transport_and_communication, finance_business_admin, educatoin_health )
```

```{r}

#save dataset selected variables
x2011_oa_var %>%
saveRDS("/home/myk9/diss/x2011_oa_var.rds")
x2011_oa_var <
readRDS("/home/myk9/diss/x2011_oa_var.rds")
```


#upload datasets 2021
#sourced: https://www.nomisweb.co.uk/sources/census_2021_bulk

x2021_oa_var <- data2021_oa %>%
#employment
mutate(employed = economic_activity_status_economically_active_excluding_full_time_students_in_employment)
mutate(unemployed = economic_activity_status_economically_active_excluding_full_time_students_unemployed
mutate(students= economic_activity_status_economically_active_and_a_full_time_student+
economic_activity_status_economically_inactive_student) %>%
rename(retired=economic_activity_status_economically_inactive_retired) %>%
#select(unemployed,employed,students,retired) %>%
#ethnicity
rename(white=ethnic_group_white) %>%
rename(mixed=ethnic_group_mixed_or_multiple_ethnic_groups) %>%
rename(indian=ethnic_group_asian_asian_british_or_asian_welsh_indian) %>%
rename(pakisthani=ethnic_group_asian_asian_british_or_asian_welsh_pakistani) %>%
rename(bangladeshi=ethnic_group_asian_asian_british_or_asian_welsh_bangladeshi) %>%
rename(chinese=ethnic_group_asian_asian_british_or_asian_welsh_chinese ) %>%
rename(black_african=ethnic_group_black_black_british_black_welsh_caribbean_or_african) %>%
rename(black_caribbean= ethnic_group_mixed_or_multiple_ethnic_groups_white_and_black_caribbean) %>%
mutate(other_ethinicity=ethnic_group_asian_asian_british_or_asian_welsh_other_asian+ethnic_group_other_ethnic_group)
#select(white, mixed,indian,pakisthani,bangladeshi,chinese,black,other_ethinicity) %>%
#age
mutate('0_to_19' = age_aged_4_years_and_under + age_aged_5_to_9_years + age_aged_10_to_14_years+age_aged_15_to_19_years)
mutate('20_to_29'= age_aged_20_to_24_years + age_aged_25_to_29_years) %>%
mutate('30_to_44' =age_aged_30_to_34_years+age_aged_35_to_39_years+age_aged_40_to_44_years) %>%
mutate('45_to_64'= age_aged_45_to_49_years + age_aged_50_to_54_years+age_aged_55_to_59_years+age_aged_60_to_64_years)
mutate(over65= age_aged_65_to_69_years+age_aged_70_to_74_years+age_aged_75_to_79_years+age_aged_80_to_84_years
#select('0_to_14','15_to_29','30_to_44','45_to_64',over65) %>%
#country of birth
rename(uk=country_of_birth_europe_united_kingdom_measures_value) %>%
mutate('other_erouope'=country_of_birth_europe_eu_countries_measures_value+
country_of_birth_europe_non_eu_countries_measures_value) %>%
rename(africa=country_of_birth_africa_measures_value) %>%
rename(asia=country_of_birth_middle_east_and_asia_measures_value) %>%
rename(america=country_of_birth_the_americas_and_the_caribbean_measures_value) %>%
rename(ocenia= country_of_birth_antarctica_and_oceania_including_australasia_and_other_measures_value)
#select(uk,other_erouope, africa, asia, america, ocenia) %>%

#tenure
mutate(owned= tenure_of_household_owned) %>%
rename('prived_rented'= tenure_of_household_private_rented) %>%
rename('social_rented'=tenure_of_household_social_rented) %>%
#select(owned,prived_rented,social_rented) %>%
#mode travel
mutate('work_home'=method_of_travel_to_workplace_work_mainly_at_or_from_home) %>%
mutate('public_transport'=method_of_travel_to_workplace_underground_metro_light_rail_tram+method_of_travel_to_workplace_train
mutate('private_transport'=method_of_travel_to_workplace_taxi+method_of_travel_to_workplace_driving_a_car_or_van
mutate('close_distance'=method_of_travel_to_workplace_motorcycle_scooter_or_moped+method_of_travel_to_workplace_bicycle
rename('other_mode'=method_of_travel_to_workplace_other_method_of_travel_to_work) %>%
#select(work_home,public_transport,private_transport,close_distance,other_mode) %>%
#occupation
mutate('managerial_and_professional' = occupation_current_1_managers_directors_and_senior_officials +
mutate('skilled_non_manual' = occupation_current_3_associate_professional_and_technical_occupations +
rename('skilled_manual' = occupation_current_5_skilled_trades_occupations) %>%
mutate('partly_skilled' =occupation_current_6_caring_leisure_and_other_service_occupations+occupation_current_7_sales_and_customer_service_occupations
#select(managerial_and_professional,skilled_non_manual,skilled_manual,partly_skilled) %>%
#education
rename('no_qualification'= highest_level_of_qualification_no_qualifications) %>%
rename('level_1'=highest_level_of_qualification_level_1_and_entry_level_qualifications) %>%
rename('level_2'=highest_level_of_qualification_level_2_qualifications) %>%
mutate('level_3'=highest_level_of_qualification_apprenticeship+highest_level_of_qualification_level_3_qualifications)
mutate('level_4'= highest_level_of_qualification_level_4_qualifications_and_above+highest_level_of_qualification_other_qualifications)
#select(no_qualification,level_1,level_2,level_3,level_4) %>%
#Dwelling
rename(detached= accommodation_type_detached) %>%
rename('semi_detached'= accommodation_type_semi_detached) %>%
rename(terrace= accommodation_type_terraced ) %>%
rename(flats=accommodation_type_in_a_purpose_built_block_of_flats_or_tenement) %>%
mutate('other_accomodation'=accommodation_type_a_caravan_or_other_mobile_or_temporary_structure) %>%
#select(detached,semi_detached,terrace,flats,other_accomodation) %>%
#work_hours
rename('0-15'= hours_worked_part_time_15_hours_or_less_worked) %>%
rename('16-30'=hours_worked_part_time_16_to_30_hours_worked) %>%
rename('31-48'=hours_worked_full_time_31_to_48_hours_worked) %>%
rename(above48=hours_worked_full_time_49_or_more_hours_worked ) %>%
#select(0-15,16-30,31-48,above48) %>%
#migration
rename('migrate_within_uk'= migrant_indicator_migrant_from_within_the_uk_address_one_year_ago_was_in_the_uk_measures_value)
rename('migrate_outside_uk'=migrant_indicator_migrant_from_outside_the_uk_address_one_year_ago_was_outside_the_uk_measures_value)
#central heating
rename(house_with_central_heating= does_have_central_heating) %>%
rename(house_without_central_heating= does_not_have_central_heating) %>%

#household composition
rename('one_person' = one_person_household_y) %>%
rename('family_without_children' = single_family_household_married_or_civil_partnership_couple_no_children_x )
mutate('family_non_depended_children'= single_family_household_married_or_civil_partnership_couple_all_children_non_dependent_y
single_family_household_cohabiting_couple_family_all_children_non_dependent_y +
single_family_household_lone_parent_family_all_children_non_dependent_y) %>%
mutate('family_depended_children'= single_family_household_married_or_civil_partnership_couple_dependent_children_x
single_family_household_lone_parent_family_with_dependent_children_y+single_family_household_cohabiting_couple_family_with_dependent_children_y
other_household_types_with_dependent_children_y) %>%
#gender
rename(male= male_x) %>%
rename ( female= female_x ) %>%
#industry
rename('agriculture_energy_and_water'= a_b_d_e_agriculture_energy_and_water) %>%
rename('manufacturing'= c_manufacturing ) %>%
rename(construction= f_construction) %>%
rename(distribution_hotels_and_restaurants= g_i_distribution_hotels_and_restaurants) %>%
rename(transport_and_communication = h_j_transport_and_communication) %>%
rename(finance_business_admin= k_l_m_n_financial_real_estate_professional_and_administrative_activities)
rename( educatoin_health= o_p_q_public_administration_education_and_health ) %>%
#rename colunm for rate calculation purpose
rename(geo_code=x2021_output_area) %>%
rename(total_pop=all_persons_x) %>%
rename(total_employment= total_x) %>%
rename(total_residents_over15= economic_activity_status_total_all_usual_residents_aged_16_years_and_over)
rename(total_household= total_all_households_y) %>%
mutate(migrants= migrant_indicator_total_all_usual_residents_measures_value
migrant_indicator_address_one_year_ago_is_the_same_as_the_address_of_enumeration_measures_value)
rename(total_dwelling=accommodation_type_total_all_households) %>%
dplyr::select(geo_code,total_pop, total_employment,total_residents_over15, total_household,total_dwelling, migrants,
unemployed,employed,students,retired, white, mixed,indian,pakisthani,bangladeshi,chinese,black_african,black_caribbean,
'0_to_19','20_to_29','30_to_44','45_to_64',over65, owned,prived_rented,social_rented,
work_home,public_transport,private_transport,close_distance,other_mode,
managerial_and_professional,skilled_non_manual,skilled_manual,partly_skilled,no_qualification,level_1,level_2,level_3,level_4 ,
detached,semi_detached,terrace,flats,other_accomodation,'0-15','16-30','31-48',above48, uk,other_erouope, africa, asia, america,
ocenia, migrate_within_uk,migrate_outside_uk, house_with_central_heating, house_without_central_heating, one_person, family_without_children,
family_depended_children, family_non_depended_children, male, female, agriculture_energy_and_water, manufacturing,construction,
distribution_hotels_and_restaurants, transport_and_communication, finance_business_admin, educatoin_health )
#save dataset
x2021_oa_var %>%
saveRDS("/home/myk9/diss/x2021_oa_var.rds")


x2021_oa_var <
readRDS("/home/myk9/diss/x2021_oa_var.rds")
#rate calculation
#2001
#to avoid misscalculate rate calculation 2001 output area macth to 2021 output area first
#join 2011_2021 output look up table to 2011 selected variable dataset ; match to 2021 output areas
x2001_data_11oa <- x2001_oa_var %>%
inner_join(x2001_2011_lookup,join_by(geo_code== OA01CD))
#summarize OA11 output area
x2001_data_sum_11oa <- x2001_data_11oa %>%
dplyr:: select(OA11CD,total_pop :educatoin_health) %>%
group_by(OA11CD) %>%
summarise(across(everything(), sum))
#match 2001_2011 dataset to 2021 daraset
x2001_21_data <- x2001_data_sum_11oa %>%
inner_join(x2011_2021_lookup,join_by(OA11CD== OA11CD)) %>%
dplyr:: select(OA21CD,total_pop:educatoin_health )
#rate calculation
#rate calculation for selected variables in:economic activity, travel_ mode, occupation, industry, education level, working hours
x2001_oa_percentage <- x2001_21_data %>%
mutate(
across(
(c(work_home: other_mode, managerial_and_professional:partly_skilled,'0-15': above48, agriculture_energy_and_water
function(x){ (x /total_industry) * 100 }
)) %>%
#rate calculation for employment , umemployment and education qualifications
mutate(
across(
(c(unemployed:retired,no_qualification:level_4 )),
function(x){ (x /total_16_17age) * 100 }
)) %>%
#rate calculation for selected variables in:country of birth, age, migration, ethnic group, gender
mutate(
across(
(c(uk: ocenia, '0_to_19':over65, white: black_caribbean, male,female)),
function(x){ (x /total_pop) * 100 }
)) %>%
#rate calculation for migrants
mutate(
across(
(c(migrate_within_uk,migrate_outside_uk)),
function(x){ (x /total_pop) * 100 }
)) %>%
#rate calculation for selected variables in:Dewlling, tenure, with central heting, without central heating, household composition
mutate(
across(
(c( owned:social_rented , house_with_central_heating,house_without_central_heating,one_person: family_non_depended_children )),
function(x){ (x /total_accomadation) * 100 }
)) %>%
mutate(
across(
(c( detached:other_accomodation )),
function(x){ (x /total_dwelling) * 100 }
))
#2011
#rate calculation- 2011
#selected variables in ; mode of transportation, industry, born country, migration, gender, central heating and household composition converted to percentages and rest of
#variables are taken as percentages directly from the table
#rate calculation for mode of travel to work and industry working
x2011_oa_percentage <- x2011_oa_var %>%
mutate(
across(
(c(work_home:other_mode, agriculture_energy_and_water: educatoin_health )),
function(x){ (x / total_employment) * 100 }
)) %>%
#rate calculation for migrants
mutate(
across(
(c( migrate_within_uk,migrate_outside_uk)),
function(x){ (x / total_pop) * 100 }
)) %>%
#rate calculation for born counntry, gender,
mutate(
across(
(c(uk:ocenia, male, female)),
function(x){ (x / total_pop) * 100 }
)) %>%

#central heating, tenure, household composition
mutate(
across(
(c(house_with_central_heating: family_non_depended_children )),
function(x){ (x /total_household ) * 100 }
)) %>%
dplyr:: select(geo_code, unemployed: educatoin_health )
#rate calculation- 2021
#rate calculation for selected variables in:economic activity, travel_ mode, occupation, industry, education level, working hours
x2021_oa_percentage <- x2021_oa_var %>%
mutate(
across(
(c( work_home: other_mode, managerial_and_professional:partly_skilled,'0-15': above48, agriculture_energy_and_water
function(x){ (x /total_employment) * 100 } #total_industry
)) %>%
#rate calculation for employmet , umemployment and education qualifications
mutate(
across(
(c(unemployed:retired,no_qualification:level_4 )),
function(x){ (x /total_residents_over15) * 100 }
)) %>%
#rate calculation for selected variables in:country of birth, age, migration, ethnic group, gender
mutate(
across(
(c(uk: ocenia, '0_to_19':over65, white: black_caribbean, male,female)),
function(x){ (x /total_pop) * 100 }
)) %>%
#rate calculation for migrants
mutate(
across(
(c( migrate_within_uk,migrate_outside_uk)),
function(x){ (x /total_pop) * 100 }
)) %>%
#rate calculation for selected variables in:Dewlling, tenure, with central heting, without central heating, household composition
mutate(
across(
(c( detached:other_accomodation, owned:social_rented , house_with_central_heating,house_without_central_heating,one_person
function(x){ (x /total_household) * 100 }
))

x2021_oa_percentage <- x2021_oa_percentage %>%
dplyr:: select(geo_code, unemployed: educatoin_health)
#save dataset 2001
x2001_oa_percentage %>% saveRDS("/home/myk9/diss/x2001_oa_percentage.rds")
x2001_oa_percentage <- readRDS("/home/myk9/diss/x2001_oa_percentage.rds")
#save dataset 2011
x2011_oa_percentage %>% saveRDS("/home/myk9/diss/x2011_oa_percentage.rds")
x2011_oa_percentage <- readRDS("/home/myk9/diss/x2011_oa_percentage.rds")
#save darasets 2021
x2021_oa_percentage %>% saveRDS("/home/myk9/diss/x2021_oa_percentage.rds")
x2021_oa_percentage <- readRDS("/home/myk9/diss/x2021_oa_percentage.rds")
############################
#join 2011_2021 look up table; 2011 oa macth to 2021 OA
x2011_data <- x2011_oa_percentage %>%
inner_join(x2011_2021_lookup,join_by(geo_code== OA11CD)) %>%
dplyr::select(OA21CD,unemployed:educatoin_health )
#rename 2021 output area colunm
x2021_data <- x2021_oa_percentage %>%
rename(OA21CD=geo_code)
#select data 2001
x2001_data <- x2001_oa_percentage %>%
dplyr:: select(OA21CD , unemployed: educatoin_health)
#save data
x2001_data %>% saveRDS("/home/myk9/diss/x2001_data.rds")
x2001_data <- readRDS("/home/myk9/diss/x2001_data.rds")
#save dataset 2011
x2011_data %>% saveRDS("/home/myk9/diss/x2011_data.rds")
x2011_data <- readRDS("/home/myk9/diss/x2011_data.rds")
#save darasets 2021
x2021_data %>% saveRDS("/home/myk9/diss/x2021_data.rds")
x2021_data <- readRDS("/home/myk9/diss/x2021_data.rds")

###combine dataset 2001, 2011, 2021

x2001_data <- x2001_data %>% mutate(year='2001')
x2011_data <- x2011_data %>% mutate(year='2011')
x2021_data <- x2021_data %>% mutate(year='2021')
var_all_census <- rbind(x2001_data, x2011_data, x2021_data)
#save combine datasets 2001, 2011, 2021
var_all_census %>% saveRDS("/home/myk9/diss/var_all_census.rds")
var_all_census <- readRDS("/home/myk9/diss/var_all_census.rds")
#check sumamry of dataset var_all_census
summary(var_all_census)

facet_wrap(~variable,scales="free") +
theme_bw()+
labs(title="HistogramPlotfor 2001,2011,2021")
#createcorelattionplot
corrplot_matrix<-var_all_census %>%
select("unemployed":"educatoin_health") %>%
#convertingthetableintomatrix
cor(method="kendall") %>%
#plottingthecorrplot
corrplot::corrplot(method= 'color',addCoef.col= 'black',tl.cex=0.4,
col=COL2('BrBG'),number.cex=0.3,number.digits=1,
sig.level=0.01,is.corr=TRUE,number.font=0.5)

var_all_census_clean <- readRDS("/home/myk9/diss/var_all_census_clean.rds")


var_all_census_clean <- readRDS("~/diss/var_all_census_clean.rds")
#IHS foe skewed variables
var_all_census_clean$indian <- asinh(var_all_census_clean$indian)
var_all_census_clean$pakisthani <- asinh(var_all_census_clean$pakisthani)
var_all_census_clean$black_african <- asinh(var_all_census_clean$black_african)
var_all_census_clean$migrate_outside_uk <- asinh(var_all_census_clean$migrate_outside_uk)
var_all_census_clean$other_erouope <- asinh(var_all_census_clean$other_erouope)
var_all_census_clean$students <- asinh(var_all_census_clean$students)
var_all_census_clean$`20_to_29` <- asinh(var_all_census_clean$`20_to_29`)
#normalize data after IHS
normalized_data_2<- var_all_census_clean %>%
mutate(
across(
c("unemployed":"educatoin_health"),
function(x) {
(x- min(x)) / (max(x)- min(x))
}
)
)
#####################################
data_for_clusters <- normalized_data_2 %>%
dplyr:: select( OA21CD ,year, indian, pakisthani, white,uk, detached, terrace, flats, other_erouope, one_person,
owned:social_rented, level_3, level_4, finance_business_admin,managerial_and_professional,skilled_non_manual,educatoin_health, students,
data_for_clusters <-data_for_clusters %>% rename(below_15h=`0-15`)
#test optimum number for k-means
library(magrittr)
set.seed(2011)
testing_wcss <- rep(NA, 30)
testing_silhouette <- rep(NA, 30)
data_for_testing <-data_for_clusters %>%
select(-OA21CD,-year)

for(testing_kin2:30){
#Calculatekmeans
kmeans_result<
stats::kmeans(data_for_testing,centers=testing_k,iter.max=10000)
#Extract WCSS
#andsaveitinthevector
testing_wcss[testing_k]<-kmeans_result %$%tot.withinss
#Calculateaveragesilhouette
#andsaveitinthevector
testing_silhouette[testing_k]<
kmeans_result %$%cluster %>%
cluster::silhouette(
data_for_testing%>%dist()
) %>%
magrittr::extract(,3)%>%mean()
}
#Calculatethegapstatisticusingbootstrapping
testing_gap<
cluster::clusGap(
data_for_testing,
FUN=kmeans,
K.max=30, #maxnumberofclusters
B=30 #numberofsamples
)

#Plots
#testingwcss
plot(2:25,testing_wcss[2:25],type="b",xlab="NumberofClusters",
ylab="WCSS",xlim=c(1,25)) +
abline(v=6,col="red") +
abline(v=11,col="red")

#testingsilhouette
plot(2:25,testing_silhouette[2:25], type="b",xlab="NumberofClusters",
ylab="Silhouette",xlim=c(1,25)) +
abline(v=6,col="red") +
abline(v=11,col="red")

#testinggap
plot(2:25,testing_gap[["Tab"]][2:25,"gap"],type="b",xlab="NumberofClusters",
ylab="Gap",xlim=c(1,25)) +
abline(v=6,col="red") +
abline(v=11,col="red")

set.seed(2021)
# Removing the variables that were excluded from the final variables list:
clusters_kmeans <- data_for_clusters %>%
dplyr:: select(indian:below_15h) %>%
stats::kmeans(
# 5 clusters is selected for k-means
centers = 8,
iter.max = 10000
)
# Adding the demographics clusters to the table
data_for_clusters <
data_for_clusters %>%
tibble::add_column(
Geodemographic_cluster =
clusters_kmeans %$%
cluster %>%
as.character() )
Geodemographic_cluster_avg <- data_for_clusters %>%
group_by(Geodemographic_cluster) %>%
dplyr::summarise(
dplyr::across(
indian:below_15h,
mean
)
)
Geodemographic_cluster_avg <- Geodemographic_cluster_avg %>%
dplyr::mutate(Geodemographic_cluster = if_else(.$Geodemographic_cluster == 1, 'B1',
if_else(.$Geodemographic_cluster == 2, 'A2',
if_else(.$Geodemographic_cluster == 3, 'D1',
if_else(.$Geodemographic_cluster == 4,
if_else(.$Geodemographic_cluster
if_else(.$Geodemographic_cluster
if_else(.$Geodemographic_cluster
))))))))
#create heat map
heatmap_plot <- Geodemographic_cluster_avg %>%
tidyr::pivot_longer(
cols =-Geodemographic_cluster,
names_to = "clustering_dimension",
values_to = "value"
) %>%
ggplot2::ggplot(
aes(
x = clustering_dimension,
y = Geodemographic_clusteraes(
fill = value
)
) +
ggplot2::xlab("Clustering Dimension") +
ggplot2::ylab("Cluster") +
ggplot2::scale_fill_viridis_c(option = "inferno") +
ggplot2::theme_bw() +
ggplot2::theme(
axis.text.x =
element_text(
angle = 90,
vjust = 0.5,
hjust = 1
)
)

#####spider plots
# install.packages("fmsb")
library(fmsb)
cluster_label <- c("A1","A2", "B1","C1","D1","E1", "G1",'F')
par(mar = rep(2,4)) # set margins
par(mfrow = c(1,2)) # 1 row × 2 columns = 2 plots
for(cluster_number in cluster_label){
rbind (
# The radar chart requires a maximum and a minimum row
# before the actual data
rep(1, 24), # max 100% for 5 variables
rep(0, 24), # min 0% for 5 variables
Geodemographic_cluster_avg %>%
dplyr::filter(Geodemographic_cluster == cluster_number) %>%
dplyr::select(-Geodemographic_cluster ) %>%
as.data.frame()
) %>%
fmsb::radarchart(title = paste("Cluster", cluster_number ))
}

#create colors for clusters
cluster_colors <- c(
"E1" = "#FFCCCC",
"A2" = "#99CCCC",
"D1" = "#003366",
"A1" = "#CCFFCC",
"B1" = "#990033",
"G1" = "#CC0066",
"C1" = "#FFCC99" ,
'F'= "#9933Ff"
)
indian
0_to_19
students
educatoin_health
skilled_non_manual
pakisthani
below_15h
white
uk
0_to_19
students
educatoin_health
skilled_non_manual
managerial_and_professional
finance_business_admin
level_4
level_4
level_3
detached
terrace
flats
other_erouope
one_person
20_to_29
black_african
social_rented
migrate_within_uk
migrate_outside_uk
prived_rented
owned
data_for_clusters <- data_for_clusters %>%
dplyr::mutate(Geodemographic_cluster_lable = if_else(.$Geodemographic_cluster == 1, 'B1',
if_else(.$Geodemographic_cluster == 2, 'A2',
if_else(.$Geodemographic_cluster == 3, 'D1
if_else(.$Geodemographic_cluster
if_else(.$Geodemographic_cluster
if_else(.$Geodemographic_cluster
if_else(.
set.seed(2021)
library(sf)
library(mapsf)

library(dplyr)
library(tmap)
#clustervisualization
#readshapefile
leicester_map<-st_read("~/diss/Leicester/leicester.shp")

#plot2001map
map_data_2001<-leicester_map %>%filter(leicester_map$year=="2001")
mf_init(
map_data_2001,
expandBB=c(0.3,0,0,0)
)
tm_shape(map_data_2001) +
tm_fill("Geodemographic_cluster_lable",title="clusters",palette=cluster_colors) +
tm_borders() +
tm_layout(title="socio-economic clustersLeicester2001",title.position= c("right","bottom"))+
tm_credits(
"Source:OfficeforNationalStatistics,\nCensus2021.ContainsNationalStatisticsdataCrowncopyrightanddatabaseright2022;
)

#plot2011map
#2011classification_results
map_data_2011<-leicester_map%>% filter(leicester_map$year=="2011")

tm_shape(map_data_2011) +
tm_fill("Geodemographic_cluster_lable",title="clusters",palette=cluster_colors) +
tm_borders() +
tm_layout(title="socio-economic clustersLeicester2011",title.position= c("right","bottom"))+
tm_credits(
"Source:OfficeforNationalStatistics,\nCensus2021.ContainsNationalStatisticsdataCrowncopyrightanddatabaseright2022;
)

# plot 2021 map
A2
B1
C1
D1
E1
F
G1
Missing
map_data_2021 <- leicester_map %>% filter(leicester_map$year == "2021")
mf_init(
map_data_2001,
expandBB = c(0.3,0,0,0)
)
tm_shape(map_data_2021) +
tm_fill("Geodemographic_cluster_lable",title="clusters",palette=cluster_colors) +
tm_borders() +
tm_layout(title = "socio- economic clusters Leicester 2021", title.position = c("right","bottom")) +
tm_credits(
"Source: Office for National Statistics,\nCensus 2021. Contains National Statistics data Crown copyright and database right 2022;
)





#uploadcombineddataset

var_all_census_clean<-readRDS("~/diss/var_all_census_clean.rds")
#normalizationdata
normalized_data<-var_all_census_clean%>%
mutate(
across(
c("unemployed":"educatoin_health"),
function(x){
(x-min(x)) / (max(x)-min(x))
}
)
)
data_for_c_clusters<-normalized_data %>%
dplyr::select(indian,pakisthani, detached,terrace,uk,flats,white,other_erouope,one_person, `20_to_29
black_african,migrate_within_uk,migrate_outside_uk,owned:social_rented,level_3,level_4,
finance_business_admin,managerial_and_professional,skilled_non_manual,students,educatoin_health)

#calculate c-means cluster
set.seed(2021)
#select variables
data_clusters <- data_for_c_clusters %>%
dplyr::select (indian, pakisthani, detached, terrace, uk,flats,white, other_erouope, one_person, `20_to_29
black_african,migrate_within_uk, migrate_outside_uk,owned:social_rented, level_3, level_4,
finance_business_admin,managerial_and_professional,skilled_non_manual,students, '0-15',
#fuzzy c cluster implemantation
cluster_result <- cmeans(data_clusters, centers = 8, m = 1.5, iter.max = 10000)
initial_centers <- cluster_result$centers
final_centers <- t(cluster_result$centers)
#initial center
#final center
# membership matrix
cluster_membership <- as.data.frame(cluster_result$membership)
#highest mambership value cluster
cluster_result$cluster_max <- apply(cluster_result$membership, 1, which.max)
cluster_max <- as.data.frame(cluster_result$cluster_max)
#secound highest cluster membership
cluster_result$cluster_secound_max <- apply(cluster_result$membership, 1,
function(x) {
order(x, decreasing = TRUE)[2] # index of the second-highest value
})
cluster_secound_max <- as.data.frame(cluster_result$cluster_secound_max)
#fuction for cofusion matrix
co_index <- function(x) {
sorted_vals <- sort(x, decreasing = TRUE) # sort in descending order
return(sorted_vals[1]- sorted_vals[2])
}
# max- 2nd max
# Apply confusion matrix fuctin for each column
confusion_index <- apply(cluster_result$membership, 1, co_index)
#bind all datasets, membership values, highest membership and secound highest membership dataserts
data_c_clusters <- cbind(data_for_c_clusters, cluster_membership, cluster_max,cluster_secound_max,confusion_index )
#Xie and Beni index
clust=FKM(data_clusters,k=8,m=1.5,stand=1)
## Xie and Beni index
xb=XB(clust$Xca,clust$U,clust$H,clust$m)


#PC values PE values plots
set.seed(123)
# Initialize a results list
#code sourced- https://rdrr.io/cran/fclust/man/XB.html
results <- data.frame(
n_clusters = integer(),
PC = numeric(),
PE=numeric()
)
for (i in 3:10) {
n_cluster <- i
m <- 2
# Run fuzzy c-means
cluster_result <- cmeans(data_clusters, centers = n_cluster, m = m)
# Calculate Partition Coefficient
PC <- PC(cluster_result$membership)
PE <- PE(cluster_result$membership)
# Store results
results <- rbind(results, data.frame(n_clusters = n_cluster, PE,PC))
}
# Plot partition coefficient
ggplot(results, aes(x = n_clusters, y = PC)) +
geom_line(color = "blue") +
geom_point(size = 2, color = "red") +
labs(
title = "Partition Coefficient vs Number of Clusters",
x = "Number of Clusters",
y = "Partition Coefficient (PC)"
) +
theme_minimal()

# Plot partition enthrophy
ggplot(results, aes(x = n_clusters, y = PE)) +
geom_line(color = "blue") +
geom_point(size = 2, color = "red") +
labs(
title = "Partition enthropy vs Number of Clusters",
x = "Number of Clusters",
y = "Partition Coefficient (PC)"
) +
theme_minimal()


#mapsformaximumclutermebership
data_c_clusters<-data_c_clusters %>%
rename(cluster_max="cluster_result$cluster_max")
#labellingclustersinsecoudmaximumcluster
data_c_clusters<-data_c_clusters %>%
dplyr::mutate(cluster_max=if_else(.$cluster_max==1, 'B1',
if_else(.$cluster_max==2, 'A2',
if_else(.$cluster_max==3, 'A1',
if_else(.$cluster_max==4, 'D1',
if_else(.$cluster_max==5, 'E1',
if_else(.$cluster_max==6, 'C1',
if_else(.$cluster_max==7
)))))))
set.seed(123)


#joinwithyourdata(data_for_clustersshouldhave 'year')
leicester_map_fuzzy<-leicester_map %>%
inner_join(data_c_clusters, join_by(OA21CD ==OA21CD))
#plotmaps
#filter2001dataset
map_data_2001<-leicester_map_fuzzy %>%filter(leicester_map_fuzzy$year =="2001")
mf_init(
map_data_2001,
expandBB=c(0.3,0,0,0)
)


tm_shape(map_data_2001) +
tm_fill("cluster_max",title="clusters",palette="BrBG") +
tm_borders() +
tm_layout(title="soci0-economic clustersLeicester2001",title.position= c("right","bottom"))+
tm_credits(
"Source:OfficeforNationalStatistics,\nCensus2021.ContainsNationalStatisticsdataCrowncopyrightanddatabaseright2022;
)














